---
title: "Comparison analysis of the original and deconvolved data in the autologous culture"
author: "Malvina Marku"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r libs, message=FALSE, warning=FALSE}
library(stringr)
library(tidyverse)
library(DESeq2)
library(limma)
library(ggplot2)
library(ggrepel)
library(ggpubr)
```

*Load Data

```{r data, message=FALSE, warning=FALSE}
# Original autologous data
#data_auto <- read.table("../Results/Normalized_data/Autologous/unnormalized_counts.txt", 
#                        header = TRUE, row.names = 1) 
data_auto <- read.table("../Results/Normalized_data/Autologous/normalized_counts.txt", 
                        header = TRUE, row.names = 1) 
names(data_auto) <- str_replace(names(data_auto), "_auto", "_autologous")

# Monoculture data
#data_B <- read.table("../Results/Normalized_data/B_cell/unnormalized_counts.txt", 
#                     header = TRUE, row.names = 1) 
data_B <- read.table("../Results/Normalized_data/B_cell/normalized_counts.txt", 
                     header = TRUE, row.names = 1) 
names(data_B) <- str_replace(names(data_B), "_B", "_monoculture")

# Deconvolved autologous data
data_auto_deconv <- read.table("../../GRN_revision_deconvolution/output/Expression_profiles/Expression_profile_Bcells.csv", 
                               header = TRUE, row.names = 1, sep = ",") 
data_auto_deconv <- log2(data_auto_deconv + 1)
names(data_auto_deconv) <- str_replace(names(data_auto_deconv), "_auto", "_deconv")
names(data_auto_deconv) <- str_replace(names(data_auto_deconv), "J", "D")
```

**Merge Datasets
```{r data_merge, message=FALSE, warning=FALSE}
# Common genes
common_genes <- intersect(intersect(rownames(data_auto), rownames(data_B)), rownames(data_auto_deconv))

data_auto_filter <- data_auto[common_genes, , drop = FALSE] %>% drop_na()
data_B_filter <- data_B[common_genes, , drop = FALSE] %>% drop_na()
data_auto_deconv_filter <- data_auto_deconv[common_genes, , drop = FALSE] %>% drop_na()

combined_data <- cbind(data_auto_filter, data_B_filter, data_auto_deconv_filter)

combined_data_D1 <- data_auto_deconv_filter %>% dplyr::select(starts_with("D1_"))
combined_data_D4 <- data_auto_deconv_filter %>% dplyr::select(starts_with("D4_"))
combined_data_D8 <- data_auto_deconv_filter %>% dplyr::select(starts_with("D8_"))
combined_data_D11 <- data_auto_deconv_filter %>% dplyr::select(starts_with("D11_"))
combined_data_D14 <- data_auto_deconv_filter %>% dplyr::select(starts_with("D14_"))
```

*PCA Analysis
```{r PCA, message=FALSE, warning=FALSE}
sample_pca <- function(input_data, Condition){
  pca.res <- prcomp(t(input_data))
  percentVar <- round(100 * pca.res$sdev ^ 2 / sum(pca.res$sdev ^ 2), 1)
  
  dataGG <- data.frame(PC1 = pca.res$x[, 1], PC2 = pca.res$x[, 2]) %>%
    rownames_to_column("Patient") %>%
    mutate(
      Day = str_split_fixed(colnames(input_data), pattern = "_", n = 4)[, 1] %>% as.factor(),
      Patient = str_split_fixed(colnames(input_data), pattern = "_", n = 4)[, 2],
      Replicate = str_split_fixed(colnames(input_data), pattern = "_", n = 4)[, 3],
      Culture = str_split_fixed(colnames(input_data), pattern = "_", n = 4)[, 4]
    )
  
  dataGG$Day <- factor(dataGG$Day, levels = c("D1", "D4", "D8", "D11", "D14"))
  
  ggplot(dataGG, aes(PC1, PC2, col = Culture, shape = Patient)) +
    geom_point() +
    geom_text(aes(label = Day), hjust = -.25, vjust = -.25) +
    ggtitle(paste("PCA plot:", Condition, "culture")) +
    xlab(paste0("PC1, VarExp: ", percentVar[1], "%")) +
    ylab(paste0("PC2, VarExp: ", percentVar[2], "%")) +
    theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"))
}
```

```{r PCA_plot, message=FALSE, warning=FALSE}
PCA_all_conditions <- sample_pca(combined_data, "auto-, mono- and deconvoluted auto-")
PCA_all_conditions
```

*Differential Gene Expression Analysis
The analysis can be done if the un-normalised counts of the deconvoluted data are available, but it is not the case. We can perform the DGEA between the original aoutologous and monoculture, and in this case, we have to choose the un-normalised data in the previous chunk. 

```{r DGEA, message=FALSE, warning=FALSE}
sample_info = as.data.frame(stringr::str_split_fixed(colnames(data_auto_deconv_filter), pattern = "_", n = 4))
names(sample_info) <- c("Day", "Patient", "Replicate", "Condition")

ddsMat <- DESeq2::DESeqDataSetFromMatrix(countData = round(combined_data),
                                         colData = sample_info,
                                         design = ~ Condition)
dds = DESeq(ddsMat)

contrast_condition <- c("Condition", "autologous", "monoculture")
results = results(dds, contrast = contrast_condition)
results_df <- as.data.frame(results)

ggma_plot <- ggmaplot(results_df, 
                      main = "Autologous: original vs deconvoluted auto",
                      fdr = 0.05, fc = 1.5, 
                      size = 0.4,
                      palette = c("#B31B21", "#1465AC", "darkgray"),
                      genenames = rownames(results_df),
                      legend = "top", top = 20,
                      font.label = c("bold", 11),
                      font.legend = "bold",
                      font.main = "bold",
                      ggtheme = ggplot2::theme_minimal())
ggma_plot
```

To perform the DGEA on normalised data, I applied limma, following the example here 
\url{https://ucdavis-bioinformatics-training.github.io/2018-June-RNA-Seq-Workshop/thursday/DE.html}. 

```{r DGEA_limma, message=FALSE, warning=FALSE}
sample_info$Condition <- factor(sample_info$Condition)
sample_info$Day <- factor(sample_info$Day)
design <- model.matrix(~ 0+Day, data = sample_info)
colnames(design) <- levels(sample_info$Day)

v <- voom(data_auto_deconv_filter, design, plot = TRUE)

fit <- lmFit(v, design)

contrast.matrix <- makeContrasts(D4 - D1, levels = design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)

results_limma <- topTable(fit2, sort.by = "P")
results_limma_df <- results_limma %>%
  rownames_to_column("gene")

ggplot(results_limma_df, aes(x = logFC, y = -log10(P.Value))) +
  geom_point(alpha = 0.5, color = "gray") +
  geom_point(data = subset(results_limma_df, adj.P.Val < 0.05 & abs(logFC) > 1.5),
             aes(x = logFC, y = -log10(P.Value)),
             color = "red", alpha = 0.7) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "blue") +
  geom_vline(xintercept = c(-1.5, 1.5), linetype = "dashed", color = "blue") +
  geom_text_repel(data = subset(results_limma_df, adj.P.Val < 0.05 & abs(logFC) > 1.5)[1:10, ],
                  aes(label = gene),
                  size = 3, max.overlaps = 20) +
  labs(title = "Volcano plot: Original autologous vs deconv",
       x = "log2 Fold Change",
       y = "-log10(p-value)") +
  theme_minimal()

```

*Correlation Analysis
```{r corr, message=FALSE, warning=FALSE}
gene_ids <- rep(common_genes, times = ncol(data_auto_filter))
vec1 <- as.vector(as.matrix(data_auto_filter))
vec2 <- as.vector(as.matrix(data_B_filter))
vec3 <- as.vector(as.matrix(data_auto_deconv_filter))

plot_df <- data.frame(
  Gene = gene_ids,
  Dataset1 = vec1,
  Dataset2 = vec2,
  Dataset3 = vec3
)

fit <- lm(Dataset1 ~ Dataset3, data = plot_df) #choose dataset
plot_df$residuals <- abs(resid(fit))

top_resid <- plot_df[order(-plot_df$residuals), ][1:10, ]
correlation <- cor(plot_df$Dataset1, plot_df$Dataset3, method = "spearman") #choose dataset

ggplot(plot_df, aes(x = Dataset2, y = Dataset3)) + #choose dataset
  geom_point(alpha = 0.4, color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  geom_point(data = top_resid, aes(x = Dataset1, y = Dataset3), #choose dataset
             color = "darkorange", size = 2) +
  geom_text_repel(data = top_resid, aes(label = Gene),
                  color = "black", size = 3, max.overlaps = 20) +
  labs(
    title = paste("Gene expression correlation (r =", round(correlation, 3), ")"),
    x = "Original autologous",
    y = "Deconvoluted autologous"
  ) +
  theme_minimal()
```

```{r}
common_genes_prova <- intersect(rownames(data_auto), rownames(data_auto_deconv))
data_auto_pr <- data_auto[common_genes_prova, ]
data_auto_deconv_pr <- data_auto_deconv[common_genes_prova, ]

gene_correlations <- sapply(1:nrow(data_auto_pr), function(i) {
  cor(as.numeric(data_auto_pr[i, ]), as.numeric(data_auto_deconv_pr[i, ]), method = "pearson")
})
names(gene_correlations) <- rownames(data_auto_pr)

# Scatter plot with ggplot
plot(as.vector(as.matrix(data_auto_pr)),
     as.vector(as.matrix(data_auto_deconv_pr)),
     xlab = "Dataset 1 expression",
     ylab = "Dataset 2 expression",
     main = "Scatter plot of all gene expressions",
     pch = 19, cex = 0.5, col = rgb(0,0,1,0.3))

print(p)
```