---
title: "Cadot scRNAseq_CLL"
author: "Hugo Chenel, Marcelo Hurtado"
date: "2025-06-23"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Marcelo

Load RNA + ADT files
```{r}
library(data.table)
library(Seurat)
library(dplyr)

# Extended function to load RNA + ADT and return a Seurat object
load_expression_and_adt <- function(rna_path, adt_path, sample_id) {
  # Load RNA
  expr <- fread(cmd = paste("zcat", rna_path), data.table = FALSE)
  rownames(expr) <- expr[[1]]
  expr <- expr[,-1]
  expr_mat <- t(as.matrix(expr))
  seurat_obj <- CreateSeuratObject(counts = expr_mat, assay = "RNA", project = sample_id)

  # Load ADT if available
  if (!is.null(adt_path)) {
    adt <- fread(cmd = paste("zcat", adt_path), data.table = FALSE)
    rownames(adt) <- adt[[1]]
    adt <- adt[,-1]
    adt_mat <- t(as.matrix(adt))
    colnames(adt_mat) <- paste0(colnames(adt_mat), "-1") #To match barcodes of RNA
    
    # Ensure same cells
    common_cells <- intersect(colnames(seurat_obj), colnames(adt_mat))
    
    # Subset RNA and ADT object to keep only common cells
    seurat_obj <- subset(seurat_obj, cells = common_cells)
    adt_mat <- adt_mat[, colnames(adt_mat) %in% common_cells]
    
    # Ensure matching cell order
    adt_mat <- adt_mat[, colnames(seurat_obj)]
    
    # Add ADT assay
    seurat_obj[["ADT"]] <- CreateAssayObject(counts = adt_mat)
  }

  return(seurat_obj)
}

# Paths to RNA and ADT files
rna_files <- list(
  M0 = "~/Downloads/GSE152469_RAW/GSM4616298_Exp_UMI_CLL_CPT1_M0.tsv.gz",
  M3 = "~/Downloads/GSE152469_RAW/GSM4616299_Exp_UMI_CLL_CPT1_M3.tsv.gz",
  M27 = "~/Downloads/GSE152469_RAW/GSM4616300_Exp_UMI_CLL_CPT1_M27.tsv.gz"
)

adt_files <- list(
  M0 = "~/Downloads/GSE152469_RAW/GSM4616298_Ab_CLL_CPT1_M0.csv.gz",
  M3 = "~/Downloads/GSE152469_RAW/GSM4616299_Ab_CLL_CPT1_M3.csv.gz",
  M27 = "~/Downloads/GSE152469_RAW/GSM4616300_Ab_CLL_CPT1_M27.csv.gz"
)

# Load each sample with RNA + ADT
seurat_list <- lapply(names(rna_files), function(sample_id) {
  load_expression_and_adt(rna_files[[sample_id]], adt_files[[sample_id]], sample_id)
})
names(seurat_list) <- names(rna_files)
```

Create CLL seurat object
```{r}
cll_obj <- merge(seurat_list[[1]], y = seurat_list[-1])
cll_obj@meta.data$time_point = c(rep("M0", ncol(seurat_list[[1]])), rep("M3", ncol(seurat_list[[2]])), rep("M27", ncol(seurat_list[[3]]))) #Add info about time point
```

Processing scRNAseq - RNA
```{r}
# RNA
cll_obj <- NormalizeData(cll_obj)
cll_obj <- FindVariableFeatures(cll_obj)
cll_obj <- ScaleData(cll_obj)
cll_obj <- RunPCA(cll_obj)
cll_obj <- FindNeighbors(cll_obj, dims = 1:20)
cll_obj <- FindClusters(cll_obj, resolution = 0.5, verbose = FALSE)
cll_obj <- RunUMAP(cll_obj, dims = 1:20)
DimPlot(cll_obj, label = TRUE)
```

Violin plots
```{r}
DefaultAssay(cll_obj) <- "ADT"
markers = c("CD14", "CD19", "CD20", "CD279", "CD3", "CD4", "CD49d", "CD5", "CD56", "CD69", "CD8")
VlnPlot(cll_obj, features = markers, ncol = 4)
```

Cluster annotation
```{r}
DefaultAssay(cll_obj) <- "RNA"
new.cluster.ids <- c("Bcells", "Bcells", "Bcells", "Tcells", "Bcells", "Tcells", "NKcells", "Monocytes/Macrophages")
#new.cluster.ids <- c("CLL", "Bcells", "CLL", "Tcells", "Bcells", "Tcells", "NKcells", "Monocytes/Macrophages")
names(new.cluster.ids) <- levels(cll_obj)
cll_obj <- RenameIdents(cll_obj, new.cluster.ids)
cll_obj@meta.data$cell_types = as.character(Idents(cll_obj))
DimPlot(cll_obj, reduction = "umap", label = TRUE, pt.size = 0.05) 
```

Save scRNAseq object
```{r}
save(cll_obj, file = "scRNAseq_obj.RData")
```

Create cell signatures
```{r}
signatures = multideconv::create_sc_signatures(cll_obj@assays$RNA@counts, cll_obj@meta.data, "cell_types", "orig.ident", credentials.mail = "marcelo.hurtado@inserm.fr", 
                                               credentials.token = "734212f6ad77fc4eea2bdb502792f294", bulk_rna = NULL, cell_markers = NULL, name_signature = "Cadot_CLL_signature")
```

Create pseudobulk from scRNAseq
```{r}
sce_cll = Seurat::as.SingleCellExperiment(cll_obj)
aggr_counts <- glmGamPoi::pseudobulk(sce_cll, group_by = glmGamPoi::vars(orig.ident), aggregation_functions = list(counts = "rowSums2", .default = "rowSums2"))
pseudo_counts_raw = data.frame(aggr_counts@assays@data$counts) %>%
  dplyr::mutate(Sample_ref = rep(NA, nrow(.)))

pseudo_counts = ADImpute::NormalizeTPM(pseudo_counts_raw, log = F) %>%
  data.frame() %>%
  dplyr::select(-Sample_ref)

pseudo_counts_raw = pseudo_counts_raw %>%
  dplyr::select(-Sample_ref)
```

Deconvolution all without signature creation
```{r}
deconv_pseudobulk = multideconv::compute.deconvolution(pseudo_counts, normalized = F, credentials.mail = "marcelo.hurtado@inserm.fr", credentials.token = "734212f6ad77fc4eea2bdb502792f294",
                                                       file_name = "Pseudobulk", create_signature = F, workers = 6, methods =  c("CBSX", "Epidish", "DWLS"), 
                                                       signatures_exclude = c("BPRNACan","BPRNACan3DProMet", "BPRNACanProMet", "CBSX-HNSCC-scRNAseq", "CBSX-Melanoma-scRNAseq", 
                                                                              "CBSX-NSCLC-PBMCs-scRNAseq", "CCLE-TIL10", "LM22", "TIL10"))

deconv_pseudobulk = read.csv("output/Deconvolution_Pseudobulk.csv", row.names = 1)
```

Benchmarking deconvolution

Groundtruth
```{r}
tab <- table(cll_obj@meta.data$orig.ident, cll_obj@meta.data$cell_types)
cells_groundtruth <- as.data.frame.matrix(prop.table(tab, margin = 1))
colnames(cells_groundtruth) = stringr::str_replace(colnames(cells_groundtruth), "Monocytes/Macrophages", "Monocytes.Macrophages.cells")
```

Benchmark 
```{r}
source("compute_mse.R")
corr_matrix = benchmark_mse(deconv_pseudobulk, cells_groundtruth, cells_extra = c("Bcells", "NKcells", "Tcells", "Monocytes.Macrophages.cells"),
                            scatter = F, file_name = "Cadot_CLL_benchmark", plot = T, height = 8, width = 10)
```

Save results
```{r}
save(pseudo_counts_raw, pseudo_counts, deconv_pseudobulk, file = "Pseudobulk_results.RData")
```

# Hugo

## Load libraries
```{r}
library(Seurat)
library(ggplot2)
library(ggpubr)
library(msigdbr)
library(dplyr)
```

## Data exploration
```{r}
# scRNA annotated with all B cells
#seurat_obj_allB <- load(file = "Data/Cadot/scRNAseq_obj.RData")

# scRNA annotated making the distinction between normal Bcells and B-CLL cells
seurat_obj_B <- load(file = "Data/Cadot/scRNAseq_obj_CLL.RData")

# Basic infos
cll_obj

# Cell counts per cell type
table(cll_obj$cell_types)
# UMAP plot colored by cell type
DimPlot(cll_obj, group.by = "cell_types")
DimPlot(cll_obj, group.by = "cell_types", label = TRUE) + NoLegend() + ggtitle("Healthy and B-CLL cells")

## Merge Bcells and CLL into one category: "B-CLLcells"
cll_obj$cell_types <- as.character(cll_obj$cell_types)
cll_obj$cell_types[cll_obj$cell_types %in% c("Bcells", "CLL")] <- "B-CLLcells"
cll_obj$cell_types <- factor(cll_obj$cell_types)
## Check result
table(cll_obj$cell_types)
DimPlot(cll_obj, group.by = "cell_types", label = TRUE) + NoLegend() + ggtitle("B-CLL cells")

# Cell counts per time point
table(cll_obj$time_point)
# UMAP plot colored by time point
DimPlot(cll_obj, group.by = "time_point", label = TRUE) + NoLegend() + ggtitle("UMAP by time point")
```

## Get myeloid genes from msigdbr
```{r}
# Example: all human GO terms containing "monocyte"
msig_mono <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:BP") %>%
  filter(grepl("MONOCYTE", gs_name, ignore.case = TRUE))

# Check all the monocyte sets
unique_sets <- unique(msig_mono$gs_name)

# Pick one signature, e.g. GOBP_MONOCYTE_DIFFERENTIATION
myeloid_genes <- msig_mono %>%
  filter(gs_name == "GOBP_MONOCYTE_DIFFERENTIATION") %>%
  pull(gene_symbol)


diff_genes <- read.table("Data/DE_genes_auto_mono.csv", header = TRUE, sep=",")
names(diff_genes)
diff_genes_all <- as.character(diff_genes$Gene)
print(intersect(myeloid_genes, diff_genes_all))

# Assuming your main dataframe is named diff_genes

# 1. Upregulated genes in "auto" condition
diff_genes_up_auto <- subset(diff_genes, Condition == "auto" & Regulation == "Up")
# 2. Downregulated genes in "auto" condition
diff_genes_down_auto <- subset(diff_genes, Condition == "auto" & Regulation == "Down")
# 3. Upregulated genes in "B" condition
diff_genes_up_B <- subset(diff_genes, Condition == "B" & Regulation == "Up")
# 4. Downregulated genes in "B" condition
diff_genes_down_B <- subset(diff_genes, Condition == "B" & Regulation == "Down")

# Convert
diff_genes_up_auto <- as.character(diff_genes_up_auto$Gene)
diff_genes_down_auto <- as.character(diff_genes_down_auto$Gene)
diff_genes_up_B <- as.character(diff_genes_up_B$Gene)
diff_genes_down_B <- as.character(diff_genes_down_B$Gene)

subset_diff_genes_up_auto <- c("APOC1", "APOE", "C2", "CHI3L1", "CISH", "ENSG00000257607", "ENSG00000271787", "FABP3", "FOLR2", "SCN1B", "SPP1")
myeloid_genes <- subset_diff_genes_up_auto
```

## Myeloid genes exploration
```{r}
# Curated myeloid gene set (edit to your list)
# myeloid_genes <- c("LYZ", "S100A9", "CD14", "ITGAM", "FCGR3A", "S100A8", "LST1", "CTSS")

# Keep only genes present in the object to avoid warnings
myeloid_genes <- intersect(myeloid_genes, rownames(cll_obj))
if (length(myeloid_genes) < 3) {
  warning("Very few myeloid genes found in the object; consider updating the list.")
}

# Compute module score ONCE on the whole object so all subsets share the same scale
# The first list element name becomes a prefix; Seurat will create 'MyeloidScore1'
cll_obj <- AddModuleScore(
  cll_obj,
  features = list(Myeloid_Score = myeloid_genes),
  name = "MyeloidScore",
  nbin = 24,            # default; adjust if  data is very small/large
  ctrl = 100            # default control genes per bin
)

# Helper for clean violin plots
vln_clean <- function(obj, group.by, title, feature = "MyeloidScore1",
                      rotate_x = FALSE, order_levels = NULL) {
  if (!is.null(order_levels)) {
    vals <- FetchData(obj, vars = group.by)[, 1, drop = TRUE]
    vals <- factor(vals, levels = order_levels)
    names(vals) <- colnames(obj)                  # <-- ensure names match cells
    obj[[group.by]] <- vals                       # safe assignment
  }

  VlnPlot(obj, features = feature, group.by = group.by, pt.size = 0) +
    labs(title = title, x = group.by, y = "Module score") +
    theme_bw(base_size = 11) +
    theme(legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = if (rotate_x) 45 else 0,
                                     hjust = if (rotate_x) 1 else 0.5))
}
```

## Disease progression: CLL early (M0, M3) vs late (M27)
```{r}
# Subset to CLL only
cll_only <- subset(cll_obj, subset = cell_types == "B-CLLcells")

# Keep only time points of interest; order them
levels_tp <- c("M0", "M3", "M27")
cll_only <- subset(cll_only, subset = time_point %in% levels_tp)
cll_only$time_point <- factor(cll_only$time_point, levels = levels_tp)

# Plot
p_progression <- vln_clean(
  cll_only,
  group.by = "time_point",
  title = "CLL cells: Upregulated genes monoculture condition score across time (M0, M3 → M27)",
  order_levels = levels_tp
)
p_progression

p_progression + 
  stat_compare_means(comparisons = list(c("M0","M3"), c("M3","M27"), c("M0","M27")),
                     method = "wilcox.test")
p_progression + geom_boxplot(width = 0.1, fill = "white", outlier.shape = NA)

# Wilcoxon rank-sum pairwise tests
pairwise.wilcox.test(
  x = cll_only$MyeloidScore1,
  g = cll_only$time_point,
  p.adjust.method = "BH"
)

p_dot <- DotPlot(
  cll_only,
  features = myeloid_genes,
  group.by = "time_point",
  cols = c("lightgrey","red")
) + 
  RotatedAxis() +
  theme_bw(base_size = 11) +
  labs(title = "Panel Downregulated genes: CLL cells", x = "Gene", y = "Time point") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_dot  # avg exp (color) and pct expressing (dot size)

# Spatial context: UMAP FeaturePlots split by time
# Pick 3–4 genes for clear viewing
#fp_genes <- intersect(c("LYZ","S100A9","CD14","ITGAM"), myeloid_genes)
fp_genes <- intersect(c("CCL4","DDIT4","MTHFD2"), myeloid_genes)

p_feat <- FeaturePlot(
  cll_only,
  features = fp_genes,
  split.by  = "time_point",
  ncol      = length(fp_genes),
  order     = TRUE
) & theme(plot.title = element_text(hjust = 0.5))

p_feat  # one panel per gene, split by M0/M3/M27
```

## Malignant vs normal B (CLL vs B cells)
```{r}
cll_b <- subset(cll_obj, subset = cell_types %in% c("CLL", "Bcells"))
cll_b$cell_types <- factor(cll_b$cell_types, levels = c("Bcells", "CLL"))

p_cll_vs_b <- vln_clean(
  cll_b,
  group.by = "cell_types",
  title = "Module score (Upregulated genes D4 vs D1): CLL vs B cells",
  order_levels = c("Bcells", "CLL")
)
p_cll_vs_b
```
**CLL vs B cells**
What it tests: Is the (Downregulated D1→D4) signature different between malignant B-CLL and normal B cells of the same lineage?
Why it’s useful: this is your lineage-matched control—it tells you tumor vs normal within B lineage.

## Absolute reference: CLL vs Monocytes/Macrophages
```{r}
cll_mono <- subset(cll_obj, subset = cell_types %in% c("B-CLLcells", "Monocytes/Macrophages"))
cll_mono$cell_types <- factor(cll_mono$cell_types, levels = c("B-CLLcells", "Monocytes/Macrophages"))

p_cll_vs_myeloid <- vln_clean(
  cll_mono,
  group.by = "cell_types",
  title = "Module score (Upregulated genes in monoculture): CLL vs Monocytes/Macrophages (reference)",
  order_levels = c("B-CLLcells", "Monocytes/Macrophages")
)
p_cll_vs_myeloid
```
**CLL vs Monocytes/Macrophages**
What it tests: Where do CLL cells sit relative to a myeloid reference for this signature?
Why it’s useful: this is an absolute reference—are CLL scores closer to (or distinct from) a canonical myeloid population?


## UMAPs for revision
```{r}
# Compute a signature score

# Compute module scores for each gene set
cll_obj <- AddModuleScore(
  object = cll_obj,
  features = list(
    diff_genes_up_auto,
    diff_genes_down_auto,
    diff_genes_up_B,
    diff_genes_down_B,
    subset_diff_genes_up_auto
  ),
  name = c("AutoUpScore", "AutoDownScore", "BUpScore", "BDownScore", "AutoUpScoreSubset")
)

grep("Score", colnames(cll_obj@meta.data), value = TRUE)

test1 <- FeaturePlot(cll_obj, features = "AutoUpScore1", cols = c("lightgrey", "darkred"),
            pt.size = 0.5) + ggtitle("Up-regulated genes in autologous culture (Cadot scRNA-seq dataset)")

test2 <- FeaturePlot(cll_obj, features = "AutoDownScore2", cols = c("lightgrey", "darkblue"),
            pt.size = 0.5) + ggtitle("Down-regulated genes in autologous culture (Cadot scRNA-seq dataset)")

test3 <- FeaturePlot(cll_obj, features = "BUpScore3", cols = c("lightgrey", "darkgreen"),
            pt.size = 0.5) + ggtitle("Up-regulated genes in monoculture (Cadot scRNA-seq dataset)")

test4 <- FeaturePlot(cll_obj, features = "BDownScore4", cols = c("lightgrey", "purple"),
            pt.size = 0.5) + ggtitle("Down-regulated genes in monoculture (Cadot scRNA-seq dataset)")

(test1 + test2 + test3 + test4) + plot_layout(ncol = 2, nrow = 2)


# Visualize on the UMAP
FeaturePlot(
  cll_obj,
  features = "AutoUpScore1",
  min.cutoff = "q05",  # 5th percentile
  max.cutoff = "q95",  # 95th percentile
  cols = c("white", "firebrick3"),
  pt.size = 0.5
) + ggtitle("Combined auto-up gene signature")

FeaturePlot(
  cll_obj,
  features = "AutoUpScore1",
  cols = c("grey90", "deepskyblue2", "darkblue"),
  pt.size = 0.5
)
FeaturePlot(
  cll_obj,
  features = "AutoUpScoreSubset5",
  cols = c("grey90", "deepskyblue2", "darkblue"),
  pt.size = 0.5
)


# Remove extrem outliers
FeaturePlot(
  cll_obj,
  features   = "AutoUpScore1",
  pt.size    = 0.4,
  order      = TRUE,
  raster     = TRUE,
  min.cutoff = -0.01,    # just below your lower bound
  max.cutoff =  0.03     # just above where most data ends
) +
  scale_color_gradientn(colors = c("grey90", "gold", "red"),
                        limits = c(-0.01, 0.03),
                        oob = scales::squish) +
  ggtitle("AutoUpScore1 (rescaled for visible contrast)")

# Remove outliers (quartiles)
FeaturePlot(
  cll_obj,
  features   = "AutoUpScoreSubset5",
  pt.size    = 0.5,
  order      = TRUE,
  raster     = TRUE,
  min.cutoff = "q01",
  max.cutoff = "q99"
) +
  scale_color_gradientn(colors = viridis::viridis(20))+
  ggtitle("Up-regulated genes in autologous culture (Cadot scRNA-seq dataset)")



# DimPlot: cell type labels only (no legend)
p1 <- DimPlot(
  cll_obj,
  group.by = "cell_types",
  label = TRUE
) +
  NoLegend() +
  ggtitle("Cadot scRNA-seq dataset annotated")

# FeaturePlot: AutoUpScore1 signature
p2 <- FeaturePlot(
  cll_obj,
  features   = "AutoUpScore1",
  pt.size    = 0.3,
  order      = TRUE,
  raster     = TRUE,
  min.cutoff = "q01",
  max.cutoff = "q99"
) +
  scale_color_gradientn(colors = viridis::viridis(20)) +
  ggtitle("Up-regulated genes in autologous culture (Cadot scRNA-seq dataset)")

# Combine plots (requires patchwork)
p1 + p2

```

# Reviewer figure
```{r}
test1 <- FeaturePlot(cll_obj, features = "AutoUpScore1", cols = c("lightgrey", "darkred"),
            pt.size = 0.5) + ggtitle("Up-regulated genes in autologous culture (Cadot scRNA-seq dataset)")

test2 <- suppressWarnings(
  FeaturePlot(
    cll_obj,
    features = "AutoUpScore1",
    cols = c("lightgrey", "darkred"),
    pt.size = 0.5
  ) +
    scale_color_gradient(
      low = "lightgrey",
      high = "darkred",
      trans = "log2"
      #name = "log2(AutoUpScore1)"
    ) +
    ggtitle("Up-regulated genes in autologous culture (Cadot scRNA-seq dataset)")
)

cll_obj$log2_AutoUpScore1 <- log2(cll_obj$AutoUpScore1 + 1)


library(viridis)
library(ggplot2)
library(systemfonts)
library(svglite)

# Build the plot
test3 <- suppressWarnings(
  FeaturePlot(
    cll_obj,
    label = TRUE,
    label.size = 5,
    features = "AutoUpScore1",
    pt.size = 0.5
  ) +
    scale_color_viridis(
      option = "A",          # A (magma), B (inferno), C (plasma), D (viridis)
      direction = 1,
      trans = "log2",
      name = "Autologous up-regulation\nmodule score"
    ) +
    ggtitle("Up-regulated genes in autologous culture (Cadot scRNA-seq dataset)") +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      legend.title = element_text(size = 11),
      legend.text  = element_text(size = 10),
      legend.key.height = unit(0.35, "cm"),
      legend.key.width  = unit(0.35, "cm")
    )
)

# Save as a small SVG (vector)
# (svglite is used under the hood by ggsave's svg device)
ggsave(
  filename = "Results/Plot/AutoUpScore1_annot.svg",
  plot     = test3,
  device   = "svg",
  width    = 8.5,   # inches
  height   = 5.0,   # inches
  units    = "in",
  limitsize = FALSE
)
```


## Aucell
```{r}
## ---- AUCell scoring ------------------------------------------
if (!requireNamespace("AUCell", quietly = TRUE)) {
  install.packages("AUCell")
}
library(AUCell)

# Signatures
up_genes_in   <- intersect(up,   rownames(cll_obj))
down_genes_in <- intersect(down, rownames(cll_obj))

# AUCell needs raw counts (or non-negative matrix); get data from active assay
assay_use <- DefaultAssay(cll_obj)
expr_mat  <- GetAssayData(cll_obj, assay = assay_use, slot = "counts")
if (max(expr_mat) == 0) expr_mat <- GetAssayData(cll_obj, assay = assay_use, slot = "data")

# Build rankings
cells_rankings <- AUCell_buildRankings(expr_mat, nCores = 1, plotStats = FALSE, verbose = FALSE)

# Calculate AUC
geneSets <- list(Up_D4vsD1 = up_genes_in, Down_D4vsD1 = down_genes_in)
auc <- AUCell_calcAUC(geneSets, cells_rankings, aucMaxRank = ceiling(0.05 * nrow(expr_mat)))

# Add to meta.data
auc_mat <- as.data.frame(t(getAUC(auc)))
cll_obj$AUC_Up   <- auc_mat$Up_D4vsD1
cll_obj$AUC_Down <- auc_mat$Down_D4vsD1
cll_obj$AUC_Net  <- cll_obj$AUC_Up - cll_obj$AUC_Down

# Order time points
levels_tp <- c("M0","M3","M27")
cll_obj$time_point <- factor(cll_obj$time_point, levels = levels_tp)

# Plot AUC across time (CLL only, or B-CLLcells after merge)
cll_only <- subset(cll_obj, subset = cell_types == "B-CLLcells")
p_auc <- VlnPlot(
  cll_only, features = c("AUC_Up","AUC_Down","AUC_Net"),
  group.by = "time_point", pt.size = 0, combine = FALSE
)
p_auc <- lapply(p_auc, function(p) p + geom_boxplot(width = 0.12, fill = "white", outlier.shape = NA) +
                  theme_bw(base_size = 11) + theme(legend.position = "none"))
p_auc[[1]] + ggtitle("AUCell: Up signature (D4 vs D1) across M0 → M3 → M27")
p_auc[[2]] + ggtitle("AUCell: Down signature (D4 vs D1) across M0 → M3 → M27")
p_auc[[3]] + ggtitle("AUCell: Net (Up − Down) across M0 → M3 → M27")

# Stats on AUC (example: Up signature)
ggpubr::stat_compare_means(
  data = FetchData(cll_only, vars = c("time_point","AUC_Up")),
  aes(x = time_point, y = AUC_Up),
  comparisons = list(c("M0","M3"), c("M3","M27"), c("M0","M27")),
  method = "wilcox.test"
)

```

